Using physical simulations to predict children's and adults' preference for physical interactions

### 2021-01-31 analysis/model_analysis_v2.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
require(tidyverse)
require(here)
```

@@

```{r, load-data}

colMeans(mdat[,1:22])
# obj_final_position_invstd_objects=drop = -0.2980893  should that be negative?

@@

@@

```{r}
load(file=here("paper/processed_data.RData"))

@@

get_model_target_choice_props <- function(feat_df, hum_df, beta) {
  hum_df <- hum_df %>% arrange(relation, drop) # same order as model df
  hum_df$Model_targ_prop = 0
  for(i in 1:nrow(hum_df)) {
    mod_cols = which(!is.na(feat_df[i,1:ncol(feat_df)]))[3:4]
    mod_choice = softmax(feat_df[i,mod_cols], beta=beta)
    hum_df[i,]$Model_targ_prop = as.numeric(mod_choice[hum_df[i,]$target] / sum(mod_choice))
  }
  return(hum_df)
}

evalFit <- function(feat_df, hum_df, beta) {
  hum_df <- get_model_target_choice_props(feat_df, hum_df, beta)
  #r = with(hum_df, cor(chose_target, Model_targ_prop))
  mse = with(hum_df, sum((chose_target - Model_targ_prop)^2)) / nrow(hum_df)
  return(mse) # return negative cor since DEoptim minimizes
}

@@

feat_names = names(mdat)[1:(ncol(mdat)-4)]
# for each feature, want to fit beta to maximize correlation between human and model choices
#feat_df <- get_feature_df(mdat, feat_names[1])

child_trial_agg <- child_long %>% group_by(relation, drop, target) %>% # alt, 
  summarize(chose_target=mean(chose_target, na.rm=T), n=n())

adult_fits = list()
child_fits = list()
for(feat in feat_names) {
  feat_df <- get_feature_df(mdat, feat)
  #paste(feat, evalFit(feat_df, adult_trial_agg, 1))
  adult_fits[[feat]] = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=adult_trial_agg)
  child_fits[[feat]] = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=child_trial_agg)
}
```

## Now try engineered features

```{r adult-eng-feats}
feat_names2 = names(mdat_eng)[1:(ncol(mdat_eng)-3)]

for(feat in feat_names2) {
  feat_df <- get_feature_df(mdat_eng, feat)
  #paste(feat, evalFit(feat_df, adult_trial_agg, 1))
  adult_fits[[feat]] = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=adult_trial_agg)
  child_fits[[feat]] = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=child_trial_agg)
}

@@

```{r}
adult_mfit <- tibble()
child_mfit <- tibble()

for(f in names(adult_fits)) {
  adult_mfit = bind_rows(adult_mfit, c(feature=f, 
                                       beta=as.numeric(unlist(adult_fits[[f]]$optim$bestmem)), 
                                       mse=adult_fits[[f]]$optim$bestval))
  child_mfit = bind_rows(child_mfit, c(feature=f, 
                                       beta=as.numeric(unlist(child_fits[[f]]$optim$bestmem)), 
                                       mse=child_fits[[f]]$optim$bestval))
}

adult_mfit <- adult_mfit %>% 
  mutate(beta = as.numeric(beta),
         mse = as.numeric(mse)) %>%
  arrange(mse)

```{r children}
child_mfit <- child_mfit %>% 
  mutate(beta = as.numeric(beta),
         mse = as.numeric(mse)) %>%
  arrange(mse)

fit_tab <- child_mfit %>% rename(child_mse = mse, child_beta = beta) %>%
  left_join(adult_mfit %>% rename(adult_mse = mse, adult_beta = beta)) %>% 
  arrange(adult_mse)

@@

```{r}
fits <- child_mfit %>% mutate(AgeGroup = "children") %>% 
  bind_rows(adult_mfit %>% mutate(AgeGroup = "adults"))

save(mdat, mdat_eng, fits, file=here("models/model_data.RData"))
```

@@

```{r}
ad <- plot_model_preds("support_response_sharpness_C=1e-05", fits, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_sharpness_C=1e-05", fits, child_trial_agg, age="children", mdat)

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
require(tidyverse)
require(here)

FIT_MODELS = F
```

@@

```{r, load-data, echo=F}

@@

mdat <- mdat %>% left_join(mdat_eng)

@@

```{r, echo=F}
load(file=here("paper/processed_data.RData"))

@@

get_model_target_choice_props <- function(feat_df, hum_df, beta) {
  hum_df <- hum_df %>% arrange(relation, drop) # same order as model df
  hum_df$Model_targ_prop = 0
  for(i in 1:nrow(hum_df)) {
    mod_cols = which(!is.na(feat_df[i,1:ncol(feat_df)]))[3:4] # CHECK THIS
    mod_choice = softmax(feat_df[i,mod_cols], beta=beta)
    hum_df[i,]$Model_targ_prop = as.numeric(mod_choice[hum_df[i,]$target] / sum(mod_choice))
  }
  return(hum_df)
}

evalFit <- function(feat_df, hum_df, beta, cor_objective=F, relation=NA) {
  hum_df <- get_model_target_choice_props(feat_df, hum_df, beta)
  if(!is.na(relation)) {
    hum_df <- subset(hum_df, relation==relation)
  }
  r = with(hum_df, cor(chose_target, Model_targ_prop))
  mse = with(hum_df, sum((chose_target - Model_targ_prop)^2)) / nrow(hum_df) 
  if(cor_objective) mse = mse - r # also try to minimize -correlation
  return(mse) # return negative cor since DEoptim minimizes
}

@@

feat_names = setdiff(names(mdat), c("drop","target","relation","trial"))
# for each feature, want to fit beta to maximize correlation between human and model choices
#feat_df <- get_feature_df(mdat, feat_names[1])

child_trial_agg <- child_long %>% group_by(relation, drop, target) %>% # alt, 
  summarize(chose_target=mean(chose_target, na.rm=T), n=n())
```

## Fit models to all trials (MSE objective)

Find best-fitting betas per feature for children and adults, contain and support trials together.

```{r fit-models-to-all-trials-mse, echo=F, eval=FIT_MODELS}
#adult_fits = list()
#child_fits = list()
all_trials_mse <- tibble()

for(feat in feat_names) {
  feat_df <- get_feature_df(mdat, feat)
  #paste(feat, evalFit(feat_df, adult_trial_agg, 1))
  adult_fit = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=adult_trial_agg)
  child_fit = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=child_trial_agg)
  all_trials_mse <- bind_rows(all_trials_mse, c(Age = "adults", feature=feat, beta=as.numeric(unlist(adult_fit$optim$bestmem)), 
                          mse = adult_fit$optim$bestval))
  all_trials_mse <- bind_rows(all_trials_mse, c(Age = "children", feature=feat, 
                          beta = as.numeric(unlist(child_fit$optim$bestmem)), 
                          mse = child_fit$optim$bestval))
}

all_trials_mse_r <- all_trials_mse_r %>% mutate(beta = as.numeric(beta),
                                                mse = as.numeric(mse))

save(all_trials_mse, file=here("models/all_trials_mse.RData"))
```

## Fit models to all trials (MSE + r objective)

```{r fit-models-to-all-trials-mse-and-r, echo=F, eval=FIT_MODELS}
all_trials_mse_r <- tibble()

for(feat in feat_names) {
  feat_df <- get_feature_df(mdat, feat)
  #paste(feat, evalFit(feat_df, adult_trial_agg, 1))
  adult_fit = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=adult_trial_agg, cor_objective=T)
  child_fit = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=child_trial_agg, cor_objective=T)
  all_trials_mse_r <- bind_rows(all_trials_mse_r, c(Age = "adults", feature=feat, beta=as.numeric(unlist(adult_fit$optim$bestmem)), 
                          mse = adult_fit$optim$bestval))
  all_trials_mse_r <- bind_rows(all_trials_mse_r, c(Age = "children", feature=feat, 
                          beta = as.numeric(unlist(child_fit$optim$bestmem)), 
                          mse = child_fit$optim$bestval))
}

all_trials_mse_r <- all_trials_mse_r %>% mutate(beta = as.numeric(beta),
                                                mse = as.numeric(mse))

@@

## Fit separately to contain and support trials (MSE)

```{r fit-models-to-supp-cont-trials-mse, echo=F, eval=FIT_MODELS}
supp_trials_mse <- tibble()
cont_trials_mse <- tibble()

for(feat in feat_names) {
  feat_df <- get_feature_df(mdat, feat)
  #paste(feat, evalFit(feat_df, adult_trial_agg, 1))
  adult_fit = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=adult_trial_agg, relation="support")
  child_fit = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=child_trial_agg, relation="support")
  supp_trials_mse <- bind_rows(supp_trials_mse, c(Age = "adults", feature=feat, beta=as.numeric(unlist(adult_fit$optim$bestmem)), 
                          mse = adult_fit$optim$bestval))
  supp_trials_mse <- bind_rows(supp_trials_mse, c(Age = "children", feature=feat, 
                          beta = as.numeric(unlist(child_fit$optim$bestmem)), 
                          mse = child_fit$optim$bestval))
}

supp_trials_mse <- supp_trials_mse %>% mutate(beta = as.numeric(beta),
                                                mse = as.numeric(mse))

for(feat in feat_names) {
  feat_df <- get_feature_df(mdat, feat)
  adult_fit = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=adult_trial_agg, relation="contain")
  child_fit = DEoptim(evalFit, lower=.5, upper=10, DEoptim.control(reltol=.001, NP=30, itermax=20), 
                feat_df=feat_df, hum_df=child_trial_agg, relation="contain")
  cont_trials_mse <- bind_rows(cont_trials_mse, c(Age = "adults", feature=feat, beta=as.numeric(unlist(adult_fit$optim$bestmem)), 
                          mse = adult_fit$optim$bestval))
  cont_trials_mse <- bind_rows(cont_trials_mse, c(Age = "children", feature=feat, 
                          beta = as.numeric(unlist(child_fit$optim$bestmem)), 
                          mse = child_fit$optim$bestval))
}

cont_trials_mse <- cont_trials_mse %>% mutate(beta = as.numeric(beta),
                                                mse = as.numeric(mse))

save(mdat, all_trials_mse, all_trials_mse_r, 
     supp_trials_mse, cont_trials_mse, file=here("models/model_data.RData"))
```

## Fit separately to contain and support trials (MSE + r)

NOT DONE

```{r}
load(here("models/model_data.RData"))
```

## Fits to all trials

```{r, echo=F}
fit_tab <- all_trials_mse %>% pivot_wider( names_from = Age, values_from = c(beta, mse)) %>%
  arrange(mse_adults)

@@

## Fits to all trials (MSE + r)

```{r, echo=F}
all_trials_mse_r <- all_trials_mse_r %>% mutate(beta = as.numeric(beta),
                                                mse = as.numeric(mse))

fit_tab <- all_trials_mse_r %>% pivot_wider( names_from = Age, values_from = c(beta, mse)) %>%
  arrange(mse_adults)

## Fits to support trials (MSE)

```{r, echo=F}
fit_tab <- supp_trials_mse %>% pivot_wider( names_from = Age, values_from = c(beta, mse)) %>%
  arrange(mse_adults)

## Fits to support trials (MSE)

```{r, echo=F}
fit_tab <- cont_trials_mse %>% pivot_wider(names_from = Age, values_from = c(beta, mse)) %>%
  arrange(mse_adults)

@@

@@

```{r}
ad <- plot_model_preds("support_response_sharpness_C=0.0001", fits, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_sharpness_C=0.0001", fits, child_trial_agg, age="children", mdat)

@@

```{r}
ad <- plot_model_preds("support_response_linearity_r", fits, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_linearity_r", fits, child_trial_agg, age="children", mdat)

```{r}
ad <- plot_model_preds("support_sharpness_mean", fits, adult_trial_agg, age="adults", mdat_eng)
ch <- plot_model_preds("support_sharpness_mean", fits, child_trial_agg, age="children", mdat_eng)

### 2021-01-31 paper/paper2021.Rmd

We also examined the split-half reliability of children's and adults' choices by repeatedly  splitting the samples in half and testing the correlation between the halves.
Based on 100 random samples, the split-half reliability for adults is *r* = `r round(mean(sh_subj_adult), 2)` (sd=`r round(sd(sh_subj_adult), 2)`), and for children is *r* = `r round(mean(sh_subj_child), 2)` (sd=`r round(sd(sh_subj_child), 2)`).
The overall split-half reliability on the containment trials is *r* = `r round(mean(sh_subj_contain), 2)` (sd=`r round(sd(sh_subj_contain), 2)`), and for support trials is *r* = `r round(mean(sh_subj_support), 2)` (sd=`r round(sd(sh_subj_support), 2)`).
[should we separately report these for adults and children?]

<!-- Finally, we analyze how exploratory children were in their responses: across the 12 trials, what proportion of the 9 available objects did each child utilize as drop objects? What proportion of the objects were targeted? -->

Figure 4 shows participants' mean proportion of unique objects dropped as a function of age.
Children of all ages sampled approximately equal proportions of the objects for dropping--roughly 70%, which is close to the 75% that would be expected if they were selected by chance (9 unique object occurring across 12 trials).

```{r unique-drop-objects, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Proportion of unique objects selected by participants. Error bars show boostrapped 95\\% confidence intervals.", eval=F}
# 12 trials, 9 objects; random selection would be 9/12 = 75%...
uniq_s <- hum %>% group_by(SID, Age_Group) %>%
  summarise(DropObjs = n_distinct(DropChoice), N=n(), prop=DropObjs/N)

uniq <- uniq_s %>% group_by(Age_Group) %>%
  tidyboot_mean(prop)

ggplot(uniq, aes(x=Age_Group, y=mean)) + geom_point() +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  xlab("Age (years)") + ylab("Unique Objects Chosen") + ylim(0,1) + 
  langcog::scale_fill_solarized() + ggthemes::theme_few() +
  geom_hline(yintercept=.75, lty='dashed')
```

@@

```{r load-model-data}
load(here("models/model_data.RData")) # mdat, mdat_eng, fits
```
Using a physics engine and 3D models of the objects that were shown to participants, we constructed a simulation of each trial of the experiment and attempted to predict children's and adults' choices using a variety of heuristic models of the simulated physical interactions.
For each trial's two possible drop object choices, we ran 250 simulated drops, and for each drop calculated a variety of features measuring the state of the model after the drop was completed.
We selected features that we thought may provide good metrics for what people would found interesting. 
For example, we collected for each drop the mean length of time before both objects come to rest (*avg_len*), as well as the standard deviation (SD) of that time (*len_std*).
Some features were calculated separately for drop and target objects, such as: the SD of the object's final position (*obj_final_position_std_objects*) as well as the inverse of that (*obj_final_position_invstd_objects*), the mean final distance from the drop location (*avg_final_radius_objects*), the maximum final distance from the drop location (*avg_max_radius_objects*), the ... (*max_radius_std_objects*), ... (*normed_velocity_std_after_first_collision_objects*)
<!-- support_probability, support_std, support_response_linearity_r, support_response_linearity_pv  -->
<!--len_inverse_sharpe_ratio, support_response_sharpness_C=1 -->

We measure each of these features across each possible drop in the experiment 100 times, and then use the resulting metrics to generate relative preferences for each of the drop choices on each trial. 
Finally, we compare how well each of these simple heuristic models account for children's and adults' drop preferences on each trial.

# Discussion 
This study aimed to investigate what physical interactions between objects most evoke curiosity.
After finding surprising consistency in an in-person pilot study of adults, an online study designed to make possible the same support and containment relations between dropped and target objects replicated this finding of consistency: adults are interested in dropping objects into containers, and to a lesser degree on top of supporting objects.
Experiment 2 found the same pattern in children, but with greater noise -- even to the extent that younger children were at-chance at choosing the support relation on most trials.

>>>>

We also examined the split-half reliability of children's and adults' choices by repeatedly  splitting the samples in half and testing the correlation between the halves.
Based on 100 random samples, the split-half reliability for adults is *r* = `r round(mean(sh_subj_adult), 2)` (sd=`r round(sd(sh_subj_adult), 2)`), and for children is *r* = `r round(mean(sh_subj_child), 2)` (sd=`r round(sd(sh_subj_child), 2)`).
The overall split-half reliability on the containment trials is *r* = `r round(mean(sh_subj_contain), 2)` (sd=`r round(sd(sh_subj_contain), 2)`), and for support trials is *r* = `r round(mean(sh_subj_support), 2)` (sd=`r round(sd(sh_subj_support), 2)`).
<!-- [should we separately report these for adults and children?] -->

@@

```{r load-model-data}
load(here("models/model_data.RData")) # mdat, mdat_eng, fits
```
Using a physics engine and 3D models of the objects that were shown to participants, we constructed a simulation of each trial of the experiment and attempted to predict children's and adults' choices using a variety of heuristic models of the simulated physical interactions.
For each trial's two possible drop object choices, we ran 250 simulated drops, and for each drop calculated a variety of features measuring the state of the model after the drop was completed.
We selected features that we thought may provide good metrics for what people would found interesting. 
For example, we collected for each drop the mean length of time before both objects come to rest (*avg_len*), as well as the standard deviation (SD) of that time (*len_std*).
Some features were calculated separately for drop and target objects, such as: the SD of the object's final position (*obj_final_position_std_objects*) as well as the inverse of that (*obj_final_position_invstd_objects*), the mean final distance from the drop location (*avg_final_radius_objects*), the maximum final distance from the drop location (*avg_max_radius_objects*), the ... (*max_radius_std_objects*), ... (*normed_velocity_std_after_first_collision_objects*)
<!-- support_probability, support_std, support_response_linearity_r, support_response_linearity_pv  -->
<!--len_inverse_sharpe_ratio, support_response_sharpness_C=1 -->
[ToDo: describe *skillsiness*]
We measure each of these features in 250 simulations of each possible drop in the experiment, and then use the resulting metrics to generate relative preferences for each of the drop choices on each trial. 
Finally, we compare how well each of these simple heuristic models account for children's and adults' drop preferences on each trial.

```{r}
# plot the best feature for children, adults, 
```

# Discussion 
This study aimed 1) to measure the consistency of children's and adults' preferences for particular object interactions, and 2) to determine whether people's interest was predicted by particular features of the imagined physical interactions between objects.
After finding surprising consistency in an in-person pilot study of adults, an online study designed to make possible the same support and containment relations between dropped and target objects replicated this finding of consistency: adults are interested in dropping objects into containers, and to a lesser degree on top of supporting objects.
Experiment 2 found the same pattern in children, but with greater noise -- even to the extent that younger children were at-chance at choosing the support relation on most trials.
